---
title: 'Chapter 2: Distributions'
author: "Ghislain Nono Gueye, Ph.D."
date: "3/7/2019"
output: bookdown::pdf_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load useful package and import data {-}

The data used in his chapter is the `2002FemPreg.Rds` data set

```{r, warning = FALSE, message = FALSE}
library(here)
library(dplyr)
library(ggplot2)
library(forcats)
```


# Histograms

The `table()` function in `R` computes frequencies and its output is a *named vector*. It helps to convert the output into a data frame for analysis.

# Representing histograms

The `table_to_df()` function takes in a vector of values and outputs a data frame of the values and their frequencies.

```{r}
# Base R approach
table_to_df <- function(x){
	df <- as.data.frame(table(x))
	colnames(df) <- c("value", "frequency")
	df$value <- as.numeric(as.character(df$value))
	df
}

# Tidyverse approach
table_to_df <- function(x){
	tibble(value = x) %>%
		group_by(value) %>%
		summarize(frequency = n())
}
```

Let's test the function with the following vector called `values`:

```{r}
values <- c(1, 2, 2, 3, 5)
table_to_df(values)
```

Finding the frequency of a specific value can be achieved with the use of the `sum()` function as well as the equality operator (i.e. `==`). Replicating the example in the book:

```{r}
sum(values == 2)
sum(values == 4)
```

There is no real need to wrap this code around a function, but if you really need to:

```{r}
# Base R approach
find_freq <- function(x, v){
	sum(x == v)
}

# Tidy approach 
find_freq <- function(x, v){
	tibble(values = x) %>%
		filter(values == v) %>%
		nrow()
}

find_freq(values, 2)
find_freq(values, 4)
```


The Python `Values()` method used in the book is the same as the `unique()` function in `R`. The main difference is that the R function outputs sorted values.

```{r}
unique(values)
```


# Plotting histograms

*Bar charts* (`geom_bar()`) and *histograms* (`geom_histogram()`) look similar, but they have subtle differences (check out this [article](https://www.differencebetween.com/difference-between-bar-graph-and-vs-histogram/)). Even though the visualizations below are histograms in nature, they are actually bar charts because they map specific values on the x-axis to their frequencies on the y-axis. Another hint is that the bars have spaces between them. 


# NSFG variables

The following script imports the dataset in a variable called `fempreg`. Then, the dataset is filtered by live births and the output is stored in `live_births`.

```{r}
fempreg <- readRDS(here("data", "processed", "used-in-book", "2002FemPreg.Rds"))

# Base R approach
live_births <- fempreg[fempreg$outcome == 1, ]

# Tidy approach
live_births <- fempreg %>%
	filter(outcome == 1)
```

Now, I make a function, which makes a frequency table (using the `table_to_df()` function) and then makes a corresponding histogram (i.e. actually a bar chart). The function also allows the user to specify plot labels (i.e. x-axis and plot title).

```{r}
make_histogram <- function(variable, x_label, plot_title){
	
	table_to_df(variable) %>%
	
	ggplot(aes(x = value, y = frequency)) +
	geom_col(fill = "#69789A", col = "#CBD2DD") +
	theme_classic() +
	labs(x = x_label, title = plot_title)
}
```


```{r}
make_histogram(live_births$birthwgt_lb, x_label = "Pounds", plot_title = "Pound (lb) part of birth weights")
make_histogram(live_births$birthwgt_oz, x_label = "Ounces", plot_title = "Ounce (oz) part of birth weights")
make_histogram(as.integer(live_births$agepreg), x_label = "Years", plot_title = "Mother's age at the end of pregnancy")
make_histogram(live_births$prglngth, x_label = "Weeks", plot_title = "Pregnancy length in weeks")
```


# Outliers

The `min_n()` and `max_n()` functions are equivalent to the `Smallest` and `Largest` methods. They display the smallest and largent `n` values of a vector.

```{r}
min_n <- function(x, n){
	head(unique(x), n)
}

max_n <- function(x, n){
	rev(tail(unique(x), n))
}
```


The 10 smallest pregnancy length values:

```{r}
min_n(live_births$prglngth, 10)
```

The 10 largest pregnancy length values:

```{r}
max_n(live_births$prglngth, 10)
```

In order to also view the frequencies of the smallest and largest values, the `table_to_df()` function can be used in conjucntion with the `head()` and `tail()` functions.

```{r}
# 10 largest pregnancy length values and their frequencies
table_to_df(live_births$prglngth) %>%
	head(10)

# 10 largest pregnancy length values and their frequencies
table_to_df(live_births$prglngth) %>%
	tail(10)
```


The `min_freq_n()` and `max_freq_n()` functions will show the smallest and largest values of a given variable as well as their frequencies in the data set. These functions depend on the `table_to_df()` function:


# First babies

```{r}
first_or_not <- live_births %>%
	filter(prglngth >= 27 & prglngth <=46) %>%
	mutate(first_baby = factor(if_else(birthord == 1, "first", "other"))) %>% 
	group_by(prglngth, first_baby) %>%
	summarize(count = n())

ggplot(data = first_or_not, aes(x = prglngth, y = count, fill = first_baby)) +
	geom_bar(position = "dodge", stat = "identity") +
	labs(fill = "baby order") +
	theme_classic()
```


# Summarizing distributions


# Variance


# Effect size


# Reporting results


# Exercises


# Glossary